---
title: "Looking at Biopics of Hollywood"
author: "Pradeep Adhokshaja"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{Biopics tidyverse analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import

For this project, we will be using the `tidyverse` library for data analysis. This is an R package that consists of tools such as `ggplot2`(data visualization),`tidyr`(data modification), `dplyr` (data management). The data set is obtained by calling the `fivethirtyeight` package using the command `library(fivethirtyeight)`.


After calling the necessary libraries, we look at the structure of the data. The command `str()` allows us to get a brief summary of the data we are dealing with. This saves from displaying the entire dataset. A custom theme is set which will be used for all the graphs in this project.

```{r include=FALSE}
plot_theme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 10,colour = "black",hjust=0.5),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=8),
    axis.text = element_text(size=8),
    axis.title.x = element_text(hjust=1),
    axis.title.y = element_text(hjust=1),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "bold"),
    legend.text = element_text(colour = "black", face = "bold"))
}
```

```{r message=FALSE, warning=FALSE}
library(fivethirtyeight)
# tidyverse includes ggplot2, tibble, tidyr, readr, purrr, dplyr:
library(tidyverse)
library(scales)
library(stringr)
library(highcharter)

glimpse(biopics)
```


The data consists of 14 variables(columns) and 764 observations(rows). We will dive deeper into the data by asking ourselves a few questions about


1. Frequency of Biopics Across the Years
2. Data Visualization by Gender, Subject and person of color/white protaganist
3. Box Office Earnings


## Number of Biopics Released by Year


```{r fig.width=7 }
biopics %>% 
  group_by(year_release) %>% 
  summarise(n=n()) %>%
  ggplot(aes(x=year_release,y=n)) +
  geom_bar(stat = "identity") +
  plot_theme() +
  labs(x="Year",y="Number of Releases",title="Number of Releases by Year",
       subtitle="", caption="Source:FiveThirtyEight")
```

The number of biopics have been increasing.


## Actors of Color and White Actors


### % Wise Plot By Decade


```{r fig.width=7}

poc <- biopics %>% 
  mutate(time_period=cut(year_release,breaks=11,label=c("1910-1920","1920-1930","1930-1940","1940-1950","1950-1960","1960-1970","1970-1980","1980-1990","1990-2000","2000-2010","2010-2014"))) %>% group_by(time_period,person_of_color) %>%
  summarise(n=n()) %>%
  mutate(person_of_color=ifelse(person_of_color==0,"White","Person of Color")) %>% 
  mutate(n=n/sum(n))

ggplot(poc, aes(x=time_period, y=n, fill=person_of_color)) +
  geom_bar(stat="identity") + 
  plot_theme() + 
  theme(axis.text.x = element_text(vjust=1,angle=90)) +
  scale_y_continuous(labels = percent) +
  labs(x="Decade",y="Percentage Composition",title="How do the % of movies involving actors of color change by decade?",caption="Data From: FiveThirtyEight")
```

The percentage of movies that had actors of color was the highest in the period *2010-2014*. Biopics that had actors of color were non existent during the period *1910-1930*. 



### Type of Subjects covered


*White Actors*

```{r fig.width=7}
biopics %>% 
  mutate(person_of_color=ifelse(person_of_color==0,"White","Person of Color")) %>% 
  group_by(person_of_color,type_of_subject) %>% 
  summarise(n=n()) %>%
  mutate(n=(n/sum(n))*100) %>% 
  filter(person_of_color !="Person of Color") %>% 
  mutate(n=round(n,2)) %>%
  hchart("treemap", hcaes(x = type_of_subject, value = n, color = n)) %>%
  hc_title(text="Type of Subjects of Movies involving White Actors(%)")
```





*Actors of Color*


```{r fig.width=7}
biopics %>% 
  mutate(person_of_color=ifelse(person_of_color==0,"White","Person of Color")) %>%
  group_by(person_of_color,type_of_subject) %>% 
  summarise(n=n()) %>%
  mutate(n=(n/sum(n))*100) %>% 
  filter(person_of_color =="Person of Color") %>% 
  mutate(n=round(n,2)) %>%
  hchart("treemap", hcaes(x = type_of_subject, value = n, color = n)) %>%
  hc_title(text="Type of Subjects of Movies involving  Actors of Color(%)")
```



*A side by side comparison*

```{r fig.width=7,fig.height=10}
biopics %>% 
  mutate(person_of_color = ifelse(person_of_color==0,"White","Person of Color")) %>%
  group_by(person_of_color,type_of_subject) %>% 
  summarise(n=n()) %>% 
  mutate(n=n/sum(n)) %>%
  ggplot(aes(x=type_of_subject,y=n,fill=person_of_color))+
  geom_bar(stat="identity")+
  scale_y_continuous(labels=percent)+
  plot_theme()+
  facet_wrap(~person_of_color)+
  coord_flip()+
  labs(title="How do subject coverage differ in biopics involving white actors and actors of color?",caption="Data From : FiveThirtyEight",x="Type of Subject",y="%")
```


* Within biopics that had actors of color,  topics like **Athelete**,**Activist**,**Singer**,**World Leader** and **Actor** had a higher coverage than in biopics that had white actors.




### Frequency of Movies Involving White Actors and Actors of Color


```{r fig.width=7,fig.height=8}
biopics %>% 
  mutate(person_of_color=ifelse(person_of_color==0,"White Actor","Actor of Color")) %>%
  group_by(year_release,person_of_color) %>% 
  summarise(n=n()) %>%
  ggplot(aes(x=year_release,y=n,colour=person_of_color))+
  geom_line()+
  plot_theme()+
  geom_vline(xintercept=1964,linetype=2)+
  labs(x="Year",y="Number",title="Number of Biopics by White Actors and Actors of Color",
       subtitle="",
       caption="Data from FiveThirtyEight")+
  geom_text(aes(1964,0),label="Civil Rights' Act",show.legend = F,hjust=-1,angle=90,vjust=1,inherit.aes = F,size=3)+
  geom_vline(xintercept=1974,linetype=2)+
  geom_text(aes(1974,0),label="First Successful African American  themed Sitcom",show.legend = F,
            hjust=0,vjust=1,inherit.aes = F,angle=90,size=3,colour="black")
```


The vertical dashed line above indicates the year 1964, when the Civil Rights' Act was passed. The number of biopics depicting colored persons went up in the 1970's. The first successful African American themed sitcom was released in the year 1974 ("Good Times").


## Gender Wise


### Number of Biopics for each gender

```{r fig.width=7}
biopics %>% 
  group_by(year_release,subject_sex) %>% 
  summarise(n=n()) %>% 
  rename(gender=subject_sex) %>%
  ggplot(aes(x=year_release,y=n,colour=gender))+
  geom_line()+ 
  plot_theme()+
  labs(x="Year",y="Number",title="Number of Biopics By Gender",
       subtitle="",caption="Data from FiveThirtyEight")+
  scale_x_continuous(breaks = seq(1920,2014,5))+
  theme(plot.title=element_text(size=18),axis.text.x = element_text(angle=90, vjust=1))
```


###  Change in the number of movies from previous years

Before we start plotting the data, we have to replace empty values with a zero. For that purpose we modify the data using the `gather` and `dcast` functions . These functions help us  reshape the data and assign values.

```{r fig.width=7,warning=FALSE}
year_wise_female <- biopics %>% 
  group_by(subject_sex,year_release) %>% 
  summarise(n=n()) %>% 
  filter(subject_sex=="Female") %>%
  mutate(n_new=lag(n,1)) %>% 
  mutate(change=n-n_new)
year_wise_male <- biopics %>% 
  group_by(subject_sex,year_release) %>% 
  summarise(n=n()) %>%
  filter(subject_sex=="Male") %>%
  mutate(n_new=lag(n,1)) %>% 
  mutate(change=n-n_new)
year_wise <- bind_rows(year_wise_male,year_wise_female)
year_wise %>%
   ggplot(aes(x=as.numeric(year_release),y=change)) + 
  geom_line() +
  plot_theme()+
  labs(x="Year",y="Number",title="Change in Number of Biopic Movies From Previous Year",
       subtitle="", caption="Data from FiveThirtyEight")+
  facet_wrap(~subject_sex,scales = "fixed")+
  scale_x_continuous(breaks = seq(1920,2014,5)) +
  theme(plot.title=element_text(size=18),axis.text.x = element_text(angle=90, vjust=1))
  
```

The change in the number of biopics that had male subjects shot up in 2014.


### What kind of subjects do movies about male and female protaganists deal with?

```{r fig.width=7,fig.height=7}
biopics$type_of_subject <- gsub(" ","",biopics$type_of_subject)
biopics <- biopics %>% 
     mutate(type_of_subject = strsplit(as.character(type_of_subject), "/")) %>% 
     unnest(type_of_subject)
biopics$type_of_subject <- tolower(biopics$type_of_subject)
women_percent <-biopics %>% 
  filter(subject_sex=="Female") %>%  
  group_by(type_of_subject) %>% 
  summarise(n=n()) %>% 
  mutate(percent_women=n/sum(n)) 
men_percent <- biopics %>% 
  filter(subject_sex=="Male") %>% 
  group_by(type_of_subject) %>% 
  summarise(n=n()) %>% 
  mutate(percent_man=n/sum(n))

percent_overall <- full_join(women_percent,men_percent,by="type_of_subject") %>% 
  select(-c(n.x,n.y)) %>%
  mutate(percent_women = ifelse(is.na(percent_women),0,percent_women),percent_man=ifelse(is.na(percent_man),0,percent_man))

ggplot(percent_overall, aes(x = percent_women, y = percent_man, color = (percent_women - percent_man))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 1, width = 0.3, height = 0.3) +
  geom_text(aes(label = type_of_subject), check_overlap = TRUE, vjust = 1.5) +
  scale_x_continuous(labels = percent_format(),limits = c(0,0.322)) +
  scale_y_continuous(labels = percent_format(),limits=c(0,0.322)) +
  scale_color_gradient( low = "red", high = "gray75") +
  theme(legend.position="none") +
  labs(y = "Biopics Involving Men", x ="Biopics Involving Women",text=element_text(size=10),
       plot.title=element_text(hjust=0.5))+
  ggtitle("Relative Frequency of Subjects(Gender Wise)")
```


Topics that are close to the line indicate  topics that have similar frequencies in both the sets of data. These topics include *government*,*academic* and *activist*. Topics that are far from this line are topics that are found frequently in one set but not the other.For example,a larger percentage of biopics that involved women,portrayed authors. By looking at the other side, topics like *military* had a higher percentage of coverage in biopics involving men than women.More the red color of the text, the more frequently the topic appears in biopics related to men.


### Race and Gender


```{r fig.width=10,fig.height=10}
biopics %>% mutate(cuts=cut(year_release,breaks=5,label=c("1910-1930","1930-1950","1950-1970","1970-1990","1990-2010")))%>%
  mutate(subject_race=ifelse(grepl("\\bHispanic\\b",subject_race),"Hispanic",subject_race)) %>% 
  group_by(cuts,subject_sex,subject_race)%>% filter(subject_race!="")%>%
  summarise(n=n()) %>%
  mutate(n=n/sum(n)) %>%
  ggplot(aes(x=cuts, y=n*100, fill=subject_race)) +
  geom_bar(stat="identity", position="dodge",width = 0.7)+
  plot_theme()+
  ylab("%")+
  ggtitle("Gender and Race")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("Year")+
  scale_fill_manual(values = c("#24576D", "#A113E2","#000000", "#D91460",
                                 "#28AADC", "#40cc49","#F2583F",
                               "#96503F","#ffc100","#918d58","#e98000","#d2f4d2",
                               "#cdc8b1","#a87582"))+
  facet_grid(subject_sex~.)+
  theme(plot.title=element_text(size=18),axis.text.x = element_text(angle=90, vjust=1))+
  labs(caption="Data From FiveThirtyEight")
```

Most of the biopics depict White Americans followed by African Americans.

## Box Office Information

The box office information suggests how successful the movie was. In this section, we are not taking into account the factor of inflation.


### Distributions of earnings per year

```{r fig.width=12,fig.height=12}
options(scipen=999)
biopics  %>%  
  filter(box_office!="-") %>% 
  mutate(
    box_office=gsub("$","",box_office,fixed=T),
    denom=str_sub(box_office,nchar(box_office),nchar(box_office)),
    box_office=gsub("M","",box_office),
    box_office=gsub("K","",box_office),
    box_office=as.numeric(box_office),
    box_office=ifelse(denom=="M",box_office*1000000,box_office),
    box_office=ifelse(denom=="K",box_office*1000,box_office),
    box_office= box_office/1000000
    ) %>% 
  ggplot(aes(x=as.factor(year_release), y=box_office)) + 
  geom_boxplot() +
  stat_summary(fun.y="mean", geom="point", shape=23, size=2, fill="white")+
  plot_theme()+
  labs(title="Distributions of Earning Per Year",x="Year",y="In Millions",
       subtitle="Inflation not taken into account(Diamond Point Represents Mean)",caption="Data From FiveThirtyEight")+
  facet_wrap(~subject_sex,ncol=1)+
  theme(axis.text.x = element_text(angle=90,vjust=1))
```

A large number of movies earned below \$100M.As we progress through the 80's, these numbers go higher , especially in the case of biopics based on men. The highest earning  movie based on a female protaganist was released in 1964.This was the "Sound of Music" which earned approximately \$163M.


### How do the distributions of box office earnings change by gender of subject?


```{r fig.width=7}
options(scipen=999)
biopics %>%
  filter(box_office!="-") %>% 
  mutate(
    box_office=gsub("$","",box_office,fixed=T),
    denom=str_sub(box_office,nchar(box_office),nchar(box_office)),
    box_office=gsub("M","",box_office),
    box_office=gsub("K","",box_office),
    box_office=as.numeric(box_office),
    box_office=ifelse(denom=="M",box_office*1000000,box_office),
    box_office=ifelse(denom=="K",box_office*1000,box_office),
    box_office=box_office/1000000) %>%
  ggplot(aes(x=subject_sex, y=box_office,fill=subject_sex)) + 
  geom_violin(color = "grey50")+
  xlab("Box Office") + 
  ylab("Count")  +
  stat_summary(fun.y="mean", geom="point", size=2, colour="white") +
  plot_theme() + 
  theme(legend.position="none")+
  labs(x="Gender",y="In Millions",title="Distributions of Box Office Earnings",
       subtitle="Inflation Not Taken into Account",
       caption="Source: FiveThirtyEight")
```
 
The mean earning is a little higher for biopics with male subjects. The mean is higher probably due to the outliers present.
 
 
### How do the distributions of box office earnings change by race of subject?



```{r fig.width=7,fig.height=7}
biopics %>% 
  filter(box_office!="-") %>% 
  mutate(
    box_office=gsub("$","",box_office,fixed=T),
    denom=str_sub(box_office,nchar(box_office),nchar(box_office)),
    box_office=gsub("M","",box_office),
    box_office=gsub("K","",box_office),
    box_office=as.numeric(box_office),
    box_office=ifelse(denom=="M",box_office*100000,box_office),
    box_office=ifelse(denom=="K",box_office*1000,box_office),
    box_office=box_office/1000000
  ) %>%
  ggplot(aes(x=as.factor(person_of_color), y=box_office,fill=as.factor(person_of_color))) + 
  geom_violin(color = "grey50")+
  xlab("Person of Color") + 
  ylab("Count")  +
  stat_summary(fun.y="mean", geom="point", size=2, colour="white") +
  plot_theme() + 
  theme(legend.position="none")+
  labs(x="Person of Color",y="In Millions",title="Distributions of Box Office Earnings",
       subtitle="Inflation Not Taken into Account",
       caption="Source: FiveThirtyEight")
```
 
 
More number of movies with protaganists of color earned higher in the box office. This is probably why the average box office earnings are higher.



## Results and Conclusion

1. We see that there are more movies depicting non colored protaganists than colored protaganists
2. Biopics based on male characters tend to be more military and sports themed.
3. The biopics pertaining to non colored main characters shot up after 1974.
3. The highest grossing biopics based on a woman is   The Sound of Music.
4. Biopics involving colored main characters earned higher on average in the box office.

This shows that there needs to be more involvement by directors and actors in producing movies that depict important personalities that belong to other races.

