---
title: "Looking at Biopics of Hollywood"
author: "Pradeep Adhokshaja"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: |
  %\VignetteIndexEntry{Biopics tidyverse analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Import

For this project, we will be using the `tidyverse` library for data analysis. This is an R package that consists of tools such as `ggplot2`(data visualization),`tidyr`(data modification), `dplyr` (data management). The data set is obtained by calling the `fivethirtyeight` package using the command `library(fivethirtyeight)`.


After calling the necessary libraries, we look at the structure of the data. The command `str()` allows us to get a brief summary of the data we are dealing with. This saves from displaying the entire dataset. A custom theme is set which will be used for all the graphs in this project.



```{r message=F,warning=FALSE}
library(tidyverse)
library(ggthemes)
library(scales)
library(stringr)
library(fivethirtyeight)
str(biopics)
plotTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 10,colour = "black",hjust=0.5),
    plot.subtitle = element_text(face="italic"),
    plot.caption = element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),
    panel.grid.major = element_line("grey80", size = 0.1),
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "grey80", color = "white"),
    strip.text = element_text(size=12),
    axis.title = element_text(size=8),
    axis.text = element_text(size=8),
    axis.title.x = element_text(hjust=1),
    axis.title.y = element_text(hjust=1),
    plot.background = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(colour = "black", face = "bold"),
    legend.text = element_text(colour = "black", face = "bold"))
}


```


The data consists of 14 variables(columns) and 764 observations(rows). We will dive deeper into the data by asking ourselves a few questions about


1. Frequency of Biopics Across the Years
1. Data Visualization by Gender, Subject and Colored/Non-Colored protaganist
1. Box Office Earnings


## Number of Biopics Released by Year


```{r fig.width=7 }

biopics %>% group_by(year_release) %>% summarise(n=n()) %>%
  ggplot(aes(x=year_release,y=n))+geom_bar(stat = "identity")+plotTheme()+
  labs(x="Year",y="Number of Releases",title="Number of Releases by Year",
       subtitle="",
       caption="Source:FiveThirtyEight")



```
The number of biopics have been increasing.


## Colored and Non Colored 


### % Wise Plot 


```{r fig.width=7}

poc <- biopics %>% group_by(year_release,person_of_color) %>% summarise(n=n()) %>%
  mutate(person_of_color=ifelse(person_of_color==0,"Not Person of Color","Person of Color")) %>% mutate(n=n/sum(n))

ggplot(data = poc, aes(x = year_release, y = n*100, fill = person_of_color)) +
  geom_bar(data = subset(poc, person_of_color=="Not Person of Color"),
           stat = "identity") +
  geom_bar(data = subset(poc, person_of_color=="Person of Color"),
           stat = "identity",
           position = "identity",
           mapping = aes(y = -n*100)) +
  scale_y_continuous(labels = abs) +
  labs(x="Year",y="%",title="Bar Plot Visualization",
       subtitle="% of Biopics that Feature Colored/Non Colored Characters",
       caption="Data from FiveThirtyEight")+coord_flip()+plotTheme()
  


```




### Type of Subjects covered


__Non Colored__


```{r fig.width=7}
library(highcharter)
biopics %>% mutate(person_of_color=ifelse(person_of_color==0,"Not Person of Color","Person of Color")) %>% group_by(person_of_color,type_of_subject) %>% 
  summarise(n=n()) %>%
  mutate(n=(n/sum(n))*100) %>% filter(person_of_color !="Person of Color") %>% mutate(n=round(n,2)) %>%
  hchart("treemap", hcaes(x = type_of_subject, value = n, color = n)) %>% hc_title(text="Type of Subjects of Movies involving Non Colored Actor(%)")
  

```





__Colored__


```{r fig.width=7}
biopics %>% mutate(person_of_color=ifelse(person_of_color==0,"Not Person of Color","Person of Color")) %>% group_by(person_of_color,type_of_subject) %>% 
  summarise(n=n()) %>%
  mutate(n=(n/sum(n))*100) %>% filter(person_of_color =="Person of Color") %>% mutate(n=round(n,2)) %>%
  hchart("treemap", hcaes(x = type_of_subject, value = n, color = n)) %>% hc_title(text="Type of Subjects of Movies involving  Colored Actor(%)")




```



### Frequency of Movies Involving Colored and Non Colored Actors


```{r fig.width=7}


biopics %>% mutate(person_of_color=ifelse(person_of_color==0,"Not Person of Color","Person of Color")) %>% group_by(year_release,person_of_color) %>% summarise(n=n()) %>%
ggplot(aes(x=year_release,y=n,colour=person_of_color))+
  geom_line()+plotTheme()+geom_vline(xintercept=1964,linetype=2)+labs(x="Year",y="Number",title="Number of Biopics By Colored/Non Colored",
       subtitle="",
       caption="Data from FiveThirtyEight")+geom_text(aes(1964,0),label="Civil Rights' Act",show.legend = F,hjust=-1,angle=90,vjust=1,inherit.aes = F)+geom_vline(xintercept=1974,linetype=2)+geom_text(aes(1974,0),label="First Black Sitcom",show.legend = F,hjust=-1,vjust=1,,angle=90,inherit.aes = F)
                     


```


The vertical dashed line above indicates the year 1964, when the Civil Rights' Act was passed. The number of biopics depicting colored persons went up in the 1970's. The first successful African American themed sitcom was released in the year 1974 ("Good Times").


## Gender Wise


### Number of Biopics for each gender

```{r fig.width=7}
biopics %>% group_by(year_release,subject_sex) %>% summarise(n=n()) %>% rename(gender=subject_sex) %>%
  ggplot(aes(x=year_release,y=n,colour=gender))+geom_line()+ plotTheme()+labs(x="Year",y="Number",title="Number of Biopics By Gender",
       subtitle="",
       caption="Data from FiveThirtyEight")+scale_x_continuous(breaks = seq(1920,2014,5))+theme(plot.title=element_text(size=18),axis.text.x = element_text(angle=90, vjust=1))
  


```


###  Change in the number of movies from previous years

Before we start plotting the data, we have to replace empty values with a zero. For that purpose we modify the data using the `gather` and `dcast` functions . These functions help us to reshape the data and assign values.



```{r fig.width=7,warning=FALSE}
library(tidyr)
library(reshape2)

year_wise <-  biopics %>% group_by(year_release,subject_sex) %>% summarise(n=n()) %>%  dcast(subject_sex~year_release,value.var="n") %>% gather(year,value,2:87) %>% mutate(value=ifelse(is.na(value),0,value)) %>%
  dcast(subject_sex~year,value.var="value") 


temp <- data.frame(gender=c("female","male"))

for(i in 3:87)
{
  
  temp<- cbind(temp,(year_wise[,i]-year_wise[,i-1]))
}

colnames(temp)<- c("subject_sex", "1927", "1929", "1930", "1933", "1934", 
"1935", "1936", "1937", "1938", "1939", "1940", "1941", "1942", 
"1943", "1944", "1945", "1946", "1947", "1948", "1949", "1950", 
"1951", "1952", "1953", "1954", "1955", "1956", "1957", "1958", 
"1959", "1960", "1961", "1962", "1963", "1964", "1965", "1966", 
"1967", "1968", "1969", "1970", "1971", "1972", "1973", "1974", 
"1975", "1976", "1977", "1978", "1979", "1980", "1981", "1982", 
"1983", "1984", "1985", "1986", "1987", "1988", "1989", "1990", 
"1991", "1992", "1993", "1994", "1995", "1996", "1997", "1998", 
"1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", 
"2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014"
)

temp %>% gather(year,value,2:86) %>%
   ggplot(aes(x=as.numeric(year),y=value))+geom_line()+ plotTheme()+
  labs(x="Year",y="Number",title="Change in Number of Biopic Movies From Previous Year",
       subtitle="",
       caption="Data from FiveThirtyEight")+
  facet_wrap(~subject_sex,scales = "fixed")+
  scale_x_continuous(breaks = seq(1920,2014,5))+
  theme(plot.title=element_text(size=18),axis.text.x = element_text(angle=90, vjust=1))
  
```



The change in the number of biopics that had male subjects shot up in 2014.


### What kind of subjects do movies about male and female protaganists deal with?


```{r fig.width=7}
biopics$type_of_subject <- gsub(" ","",biopics$type_of_subject)
biopics <- biopics %>% 
     mutate(type_of_subject = strsplit(as.character(type_of_subject), "/")) %>% 
     unnest(type_of_subject)
biopics$type_of_subject <- tolower(biopics$type_of_subject)
women_percent <-biopics   %>% filter(subject_sex=="Female") %>%  group_by(type_of_subject) %>% summarise(n=n()) %>% mutate(percent_women=n/sum(n)) 
men_percent <- biopics %>% filter(subject_sex=="Male") %>% group_by(type_of_subject) %>% summarise(n=n()) %>% mutate(percent_man=n/sum(n))

percent_overall <- full_join(women_percent,men_percent,by="type_of_subject") %>% select(-c(n.x,n.y)) %>%
  mutate(percent_women = ifelse(is.na(percent_women),0,percent_women),percent_man=ifelse(is.na(percent_man),0,percent_man))
ggplot(percent_overall, aes(x = percent_women, y = percent_man, color = abs(percent_women - percent_man))) +
  geom_abline(color = "gray40", lty = 2) +
  geom_jitter(alpha = 0.1, size = 1, width = 0.3, height = 0.3) +
  geom_text(aes(label = type_of_subject), check_overlap = TRUE, vjust = 1.5) +
  scale_x_continuous(labels = percent_format(),limits = c(0,0.322)) +
  scale_y_continuous(labels = percent_format(),limits=c(0,0.322)) +
  scale_color_gradient( low = "red", high = "gray75") +
  theme(legend.position="none") +
  labs(y = "Biopics Involving Men", x ="Biopics Involving Women",text=element_text(size=10),plot.title=element_text(hjust=0.5))+ggtitle("Relative Frequency of Subjects(Gender Wise)")
```


It is pretty obvious that a larger percentage of  biopics with female protaganists deal with actresses more than biopics with male protaganists. The subjects falling on the dotted line refer to subjects which had equal frequency in both the groups. These include subjects such as **government**,**academic**,**activist**. A larger percentage of movies with male protaganists were military themed.


### Race and Gender


```{r fig.width=7}
biopics %>% mutate(cuts=cut(year_release,breaks=5,label=c("1910-1930","1930-1950","1950-1970","1970-1990","1990-2010")))%>% group_by(cuts,subject_sex,subject_race)%>% filter(subject_race!="") %>%
  summarise(n=n()) %>%
  mutate(n=n/sum(n)) %>%
  ggplot(aes(x=cuts, y=n*100, fill=subject_race)) +
geom_bar(stat="identity", position="dodge",width = 0.7)+plotTheme()+ylab("%")+ggtitle("Gender and Race")+theme(plot.title = element_text(hjust = 0.5))+xlab("Year")+scale_fill_manual(values = c("#24576D", "#A113E2",
                                 "#000000", "#D91460",
                                 "#28AADC",
                                 "#40cc49",
                                 "#F2583F",
                                 "#96503F","#ffc100","#918d58","#e98000","#d2f4d2","#cdc8b1","#7c3838","#1fffaf","#a87582","#5b9c31"))+facet_grid(subject_sex~.)+theme(plot.title=element_text(size=18),axis.text.x = element_text(angle=90, vjust=1))+labs(caption="Data From FiveThirtyEight")
```

Most of the biopics depict White Americans followed by African Americans.


## Directors and Lead Actors(Decade Wise)





```{r message=FALSE,fig.width=7}
library(igraph)
library(ggraph)
library(gganimate)

g <-  biopics %>% mutate(time_period=cut(year_release,breaks=11,label=c("1910-1920","1920-1930","1930-1940","1940-1950","1950-1960","1960-1970","1970-1980","1980-1990","1990-2000","2000-2010","2010-2014"))) %>% group_by(director,lead_actor_actress,time_period) %>% summarise(n=n()) %>% graph_from_data_frame()

directors <- levels(as.factor(biopics$director))

V(g)$class <- rep("",gorder(g))

for(i in 1:gorder(g)){
  V(g)[i]$class <- ifelse(V(g)[i]$name %in% directors,"Directors",V(g)[i]$class)
}

for(i in 1:gorder(g)){
  V(g)[i]$class <- ifelse(V(g)[i]$class=="","Actor",V(g)[i]$class)
}

set.seed(100)
p <- ggraph(g, layout = 'kk') +
    geom_edge_link(aes(colour = (time_period))) +
    geom_node_point(aes(shape = class))+scale_edge_color_manual(values = c("#94edc7","#180228","#010d5c","#ef1c17","#908b5b","#f3b893","#3c5e34","#91a5ae","#ff33dd","#ff947b","#cd0000") )

p+ggtitle("Director and Actor Network by Decade")+theme(plot.title = element_text(hjust=0.5))+theme_graph()




```


For the above plot, we use the `ggraph` library by [Thomas Lin Pedersen](https://github.com/thomasp85/ggraph). Using the `scale_edge_color_manual()` function, we set the the colors for each time period. The `graph_from_data_frame()` function allows us to change the dataframe to an `igraph` object.

## Box Office Information

The box office information suggests how successful the movie was. In this section, we are not taking into account the factor of inflation.


### Distributions of earnings per year

```{r fig.width=7}
biopics  %>%  filter(box_office!="-") %>% mutate(box_office=gsub("$","",box_office,fixed=T)) %>%
  mutate(denom=str_sub(box_office,nchar(box_office),nchar(box_office))) %>%
  mutate(box_office=gsub("M","",box_office))%>%
  mutate(box_office=gsub("K","",box_office)) %>%
  mutate(box_office=as.numeric(box_office)) %>%
  mutate(box_office=ifelse(denom=="M",box_office*1000000,box_office*1000)) %>%
  ggplot(aes(x=as.factor(year_release), y=box_office)) + geom_boxplot() +
stat_summary(fun.y="mean", geom="point", shape=23, size=2, fill="white")+plotTheme()+
  labs(title="Distributions of Earning Per Year",x="Year",y="$",subtitle="Inflation not taken into account(Diamond Point Represents Mean)",caption="Data From FiveThirtyEight")+facet_wrap(~subject_sex,ncol=1)+theme(axis.text.x = element_text(angle=90,vjust=1))
```

A large number of movies earned below \$100M.As we progress through the 80's, these numbers go higher , especially in the case of biopics based on men. The highest earning  movie based on a female protaganist was released in 1964.This was the "Sound of Music" which earned approximately \$163M.


### How do the distributions of box office earnings change by gender of subject?


```{r fig.width=7}
biopics  %>%  filter(box_office!="-") %>% mutate(box_office=gsub("$","",box_office,fixed=T)) %>%
  mutate(denom=str_sub(box_office,nchar(box_office),nchar(box_office))) %>%
  mutate(box_office=gsub("M","",box_office))%>%
  mutate(box_office=gsub("K","",box_office)) %>%
  mutate(box_office=as.numeric(box_office)) %>%
  mutate(box_office=ifelse(denom=="M",box_office*1000000,box_office*1000)) %>%
  ggplot(aes(x=subject_sex, y=box_office,fill=subject_sex)) + 
  geom_violin(color = "grey50")+
  xlab("Box Office") + ylab("Count")  +
  stat_summary(fun.y="mean", geom="point", size=2, colour="white") +
  plotTheme() + theme(legend.position="none")+
  labs(x="Gender",y="($)",title="Distributions of Box Office Earnings",
       subtitle="Inflation Not Taken into Account",
       caption="Source: FiveThirtyEight")




```
 
The mean earning is a little higher for biopics with male subjects. The mean is higher probably due to the outliers present.
 
 
### Type of Subject



```{r fig.width=7}


biopics  %>%  filter(box_office!="-") %>% mutate(box_office=gsub("$","",box_office,fixed=T)) %>%
  mutate(denom=str_sub(box_office,nchar(box_office),nchar(box_office))) %>%
  mutate(box_office=gsub("M","",box_office))%>%
  mutate(box_office=gsub("K","",box_office)) %>%
  mutate(box_office=as.numeric(box_office)) %>%
  mutate(box_office=ifelse(denom=="M",box_office*1000000,box_office*1000)) %>%
  ggplot(aes(x=as.factor(person_of_color), y=box_office,fill=as.factor(person_of_color))) + 
  geom_violin(color = "grey50")+
  xlab("Person of Color") + ylab("Count")  +
  stat_summary(fun.y="mean", geom="point", size=2, colour="white") +
  plotTheme() + theme(legend.position="none")+
  labs(x="Person of Color",y="($)",title="Distributions of Box Office Earnings",
       subtitle="Inflation Not Taken into Account",
       caption="Source: FiveThirtyEight")



```
 
 
More number of movies with colored protaganists earned higher in the box office. This is probably why the average box office earnings are higher.



## Results and Conclusion

1. We see that there are more movies depicting non colored protaganists than colored protaganists
1. Biopics based on male characters tend to be more military and sports themed.
1. The biopics pertaining to non colored main characters shot up after 1974.
1. The highest grossing biopics based on a woman is   The Sound of Music.
1. Biopics involving colored main characters earned higher on average in the box office.

This shows that there needs to be more involvement by directors and actors in producing movies that depict important personalities that belong to other races.

