---
title: "Gender in Comic Books"
author: "Jonathan Bouchet"
date: "`r Sys.Date()`"
output: 
rmarkdown::html_vignette:
df_print: kable
vignette: |
  %\VignetteIndexEntry{Comic gender analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  This vignette is based on [538 study : Comic Books Are Still Made By Men, For Men And About Men](https://fivethirtyeight.com/features/women-in-comic-books/) study about Marvel and DC characters since ~1939 until 2014 (August 24th) and aims at investigating the features of Comic Books characters according their gender / Publisher.
  
```{r, message = FALSE, warning = FALSE}
#load packages and csv file
library(ggplot2)
library(dplyr)
library(grid)
library(fmsb)
library(fivethirtyeight)
library(knitr)
library(wordcloud)
library(gridExtra)
```

####Overview plots 

* percentage of Gender per publisher.
* raw number of characters per publisher.

```{r, warning = FALSE, message = FALSE, fig.width = 10}
#calculate raw number and percentage of each gender for each publisher
raw_number_per_publisher <- comic_characters %>% 
  group_by(publisher) %>% 
  summarise(number = n()) %>% 
  arrange(-number)

percent_gen_pub <- comic_characters %>% 
  group_by(sex, publisher) %>% 
  summarise(number = n()) %>% 
  arrange(-number) %>%
  group_by(publisher) %>% 
  mutate(countT = sum(number)) %>% 
  group_by(sex) %>% 
  mutate(
    percentage = (100*number/countT),
    label = paste0(round(percentage, 2)))

#plot percentage of each gender for each publisher
percentage_per_publisher <- ggplot(data = percent_gen_pub,aes(x = sex,y = percentage, fill = publisher)) + 
  geom_bar(width = 0.9, stat = "identity", position = 'dodge') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position = 'none') + 
  geom_text(aes(label = label), position=position_dodge(width = 0.9), vjust = -0.25,size = 2.5) + 
  scale_fill_manual(values = c("#3B9AB2","#EBCC2A")) + 
  xlab('') + 
  ylab('Percentage')

raw_number_per_publisher <- ggplot(data = raw_number_per_publisher, aes(x = publisher, y = number, fill = publisher)) + 
  geom_bar(width = 0.1, stat = "identity") + 
  coord_flip() +
  scale_fill_manual(values = c("#3B9AB2","#EBCC2A")) + 
  xlab('') + 
  ylab('') +
  theme(legend.position = 'None')

grid.arrange(percentage_per_publisher, raw_number_per_publisher, ncol=2)
```

* Marvel has more than the double of number of characters compared to DC.
* `Male` are more present in both publishers compared to `Female` on a ~2.5:1 ratio, while `LGBT` characters represent less than 1 percent.

#### Number of Characters vs. Time
```{r, message = FALSE, warning = FALSE, fig.width = 10}
#select data with no NA's for sex and date and groupby

#define list of gender per publisher
gender_list_marvel <- c("Female Characters", "Male Characters", "Genderfluid Characters","Agender Characters")
gender_list_dc <- c("Female Characters", "Male Characters", "Genderless Characters","Transgender Characters")

marvel_vs_time <- comic_characters %>% 
  filter(publisher == 'Marvel' & !is.na(month) & !is.na(sex)) %>% 
  group_by(year, month, sex) %>% 
  summarise(number = n()) %>% 
  mutate(
    sex_ordered = factor(sex, levels = gender_list_marvel), 
    month_ordered = factor(month, levels = month.name))

dc_vs_time <- comic_characters %>% 
  filter(publisher == 'DC' & month!= "Holiday" & !is.na(month) & !is.na(sex)) %>% 
  mutate(month = ifelse(month=="01","January",month)) %>% 
  group_by(year, month, sex) %>% summarise(number = n()) %>% 
  mutate(
    sex_ordered = factor(sex, levels = gender_list_dc), 
    month_ordered = factor(month, levels = month.name)) 

plot_marvel_time <- ggplot(data = marvel_vs_time, aes(year, month_ordered)) + 
  geom_tile(aes(fill = number),colour = "white") + 
  scale_fill_gradient(low = "#EBCC2A", high = "black") + 
  facet_wrap(~ sex_ordered, ncol = 4) +
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.title.y = element_blank(), axis.ticks.y =  element_blank(), legend.position = 'right', legend.title = element_blank(), legend.key.size = unit(.2, "cm")) + 
  xlim(1935,2015)

plot_dc_time <- ggplot(data = dc_vs_time, aes(year, month_ordered)) + geom_tile(aes(fill = number),colour = "white") + 
  scale_fill_gradient(low = "#3B9AB2", high = "black") + 
  facet_wrap(~ sex_ordered, ncol = 4) + 
  theme(axis.title.x = element_blank() ,axis.ticks.x = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank(), legend.position = 'right', legend.title = element_blank(), legend.key.size = unit(.2, "cm")) + 
  xlim(1935,2015)

grid.arrange(rbind(ggplotGrob(plot_marvel_time), ggplotGrob(plot_dc_time), size = "last"))
```

* Most of the characters are either `Female` or `Male` for both publishers.
* There is a gap in character’s creation (for both publishers) ~1955-1960.
* The increase in `Female` characters appear early for Marvel (~1970) while it was in the late 70’s for DC comics.
* `LGBT` characters appear in the late 70’s (DC) and in the late 60’s (Marvel).

####Characteristics per Gender, publisher

To represent the average characteristics, radarchart will be used.

* Characteristics looked at are : `id`,`align`,`eye`,`hair` and `alive`.
* the idea is to group the rows by these features and compute their mean
* later by taking the higher percentage, we can deduce a `profile` representative of each publisher.

#####example: vs `Hair` feature
```{r, message = FALSE, warning = FALSE}
#function to aggregate data by (gender, feature, publisher)
aggregateFeature <- function(current_publisher, feature){
  #empty list to keep dataframe by (gender, feature)
  currentFeature <- list()
  if(current_publisher == 'Marvel'){
    gender_list <- gender_list_marvel
  } else {
    gender_list <- gender_list_dc
  }
  
  for(i in 1:length(gender_list)){
    currentFeature[[i]] <- comic_characters %>% 
      filter(publisher == current_publisher ) %>%  
      select(hair, sex) %>% 
      na.omit() %>% 
      group_by(hair) %>% 
      filter(sex == gender_list[i]) %>% 
      summarise(number = n()) %>% 
      arrange(-number) %>% 
      mutate(countT = sum(number)) %>% 
      mutate(percentage = round(100*number/countT,1)) %>%
      select(hair, percentage, number)
    
    if(current_publisher == 'Marvel'){
      colnames(currentFeature[[i]])[2]<-'percentage_marvel'
    } else{
      colnames(currentFeature[[i]])[2]<-'percentage_dc'
    }
    
    if(feature == 'hair'){
      #strip 'hair' word for better display of the radarchart
      currentFeature[[i]]$hair <- sapply(currentFeature[[i]]$hair, function(x) gsub(' Hair','', x))}
  }
  
  names(currentFeature) <- gender_list
  return(currentFeature)
}
```

```{r, message = FALSE, warning = FALSE}
#outer join the 2 dataframes for (feature=Hair, gender=Male)
merged <- full_join(aggregateFeature('DC','hair')[[2]], aggregateFeature('Marvel','hair')[[2]], by='hair')
#set min/max percentages for the radarchart limits
min <- rep(0, length(merged$hair))
max <- rep(50, length(merged$hair))

maleHair <- data.frame(rbind(max,min,merged$percentage_dc,merged$percentage_marvel))
colnames(maleHair) <- merged$hair
row.names(maleHair) <- c('max','min','percentage_dc','percentage_marvel')
maleHair[is.na(maleHair)] <- 0
```

```{r, message = FALSE, warning = FALSE}
#cosmetics
radarchart(maleHair,                     #dataframe
           axistype=2,                   #axis type
           pcol=c("#3B9AB2", "#EBCC2A"), #color -Wes Anderson Sissou-
           plwd=1,                       #axis line width
           pty=19,                       #marker type
           plty=3,                       #line type
           cglcol="grey",                #axis line color
           cglty=2,                      #axis line type
           axislabcol="grey",            #Color of axis label
           cglwd=.6,                     #line type
           vlcex=.6,                     #font size label
           palcex=1.)                    #font size value
legend(x=1, y=1.3,                                               #position
       legend = c('DC male characters','Marvel male characters'),#labels
       bty="n",                                                  #no window
       pch=16,                                                   #marker type 
       text.col = "black",                                       #label colors 
       col=c("#3B9AB2","#EBCC2A"),                               #marker color
       cex=.8,                                                   #marker size 
       pt.cex=1)                                                 #font size
```

####Popularity
By looking at the number of appearances for each characters, we can make a visualization (`wordcloud`) representing the most populars characters.

```{r, message = FALSE, warning = FALSE, fig.width = 10}
#remove NA from appearance column, keep only the name inside parentheses
marvel_appearances <- comic_characters %>% 
  filter(publisher == 'Marvel') %>% 
  select(name, appearances) %>% na.omit() %>% 
  mutate(name = gsub(" *\\(.*?\\) *", "", name))
dc_appearances <- comic_characters %>% 
  filter(publisher == 'DC') %>% 
  select(name, appearances) %>% na.omit() %>% 
  mutate(name = gsub(" *\\(.*?\\) *", "", name))

color_marvel <- colorRampPalette(c("#EBCC2A", "black"))
color_dc <- colorRampPalette(c("#3B9AB2", "black"))

op <- par(mar = c(1, 2, 2, 1), mfrow = c(1, 2))
wordcloud(dc_appearances$name, dc_appearances$appearances, min.freq = 100, colors = color_dc(10), scale = c(1.75, 0.2))
wordcloud(marvel_appearances$name, marvel_appearances$appearances, min.freq = 250, colors = color_marvel(10), scale = c(1.5, 0.2))
```

* `Batman`(DC) and `Spider-Man`(Marvel) ends up being the characters with the most appearances. 